---
title: "Assignment_1 - chetan_kandula"
output:
  pdf_document: default
  html_document: default
  word_document: default
date: "2025-03-12"
---

***Loading data set and libraries***


```{r}
# Load required dataset library
library(datasets)

# Set CRAN repository explicitly before installing packages
options(repos = c(CRAN = "https://cloud.r-project.org/"))

# Install required packages
install.packages("mice", dependencies = TRUE)
install.packages("tinytex", dependencies = TRUE)

# Load the 'mice' library
library(mice)

# Read the dataset
dataset_cars <- read.csv("C:\\Users\\Asus\\Downloads\\cars_data_10K.csv")

# Check if TinyTeX is installed; install only if missing
if (!tinytex::is_tinytex()) {
  tinytex::install_tinytex(force = TRUE)
}

# Install ggplot2 if not installed
if (!requireNamespace("ggplot2", quietly = TRUE)) {
  install.packages("ggplot2", dependencies = TRUE)
}

# Load ggplot2 library
library(ggplot2)


```

***Exploring and knowing the dataset:***

In this part I have explored  and studied whole dataset and columns.

```{r}
summary(dataset_cars)
head(dataset_cars)
colnames(dataset_cars)
dim(dataset_cars)
```
***Cleaning Dataset:***

1.To check the data
2.Identifying Null[NA] values
3.Identifying special characters and replacing/removing from the column to convert the data in appropriate data
4.Handling Outliers.

*Cleaning each and every column step-by-step:*

column-1 : Make
```{r}

summary(dataset_cars$Make)
#finding how many times a word got repeated and finding frequency of the words
frequency(dataset_cars$Make)

#converting all the column names to lowercase:
dataset_cars$Make <- tolower(dataset_cars$Make)
#checking null values[NA]:
sum(is.na(dataset_cars$Make))
#removing unwanted spaces from this column:
dataset_cars$Make <- gsub(" ","",dataset_cars$Make)

#now removing all the special characters from the columns:
#In this columns we are having special character {'-'} -> removing '-' and replacing:
dataset_cars$Make <-gsub("-","",dataset_cars$Make)

#once again checking for other special characters in data:
#grepl("[^a-zA-Z ]",dataset_cars$Make)

table(dataset_cars$Make)

```
*SUMMARY OF COLUMN-1:MAKE*

1.The above column(Make) is entirely cleaned and preprocessed.
2.I have also removed all the special charactes and unwanted spaces in the column.
3.chevrolet made more no.of cars = 937


Column-2:Model

```{r}
#Summary of the column:
summary(dataset_cars$Model)
#checking null characters:
sum(is.na(dataset_cars$Model))
#frequency:
frequency(dataset_cars$Model)
#Converting to lowercase:
dataset_cars$Model <- tolower(dataset_cars$Model)

#so here we can see that there are special characters like ['-' , '/' , '.', "space"]:
#removing Special characters and converting the data into appropriate format:
dataset_cars$Model <- gsub(" ","",dataset_cars$Model)
dataset_cars$Model <- gsub("-","",dataset_cars$Model)
dataset_cars$Model <- gsub("/","",dataset_cars$Model)
dataset_cars$Model <- gsub("\\.","",dataset_cars$Model)


```
*Summary of column-2: Model*

1.The data is cleaned and preprocessed properly
2.In this column I found Special characters[-,/,.].
3.So i removed all the special characters and unwanted spaces and made the data 
  ready.


Column-3 : YEAR

```{R}
summary(dataset_cars$Year)
#validating data:
is.numeric(dataset_cars$Year)
#checking null values:
sum(is.na(dataset_cars$year))
#checking for any special characters in YEAR column:
specialchars3 <- gsub("[0-9 ]","",dataset_cars$Year)
cat(specialchars3)
#outliers
IQR_0 <- 2016 - 2007
LB_0 <- 2007 - 1.5 * IQR_0
UB_0 <- 2016 + 1.5 * IQR_0
#finding outliers and represemting them using boxplot
outliers_0 <- dataset_cars$YEAR[dataset_cars$YEAR < LB_0 | 
                            dataset_cars$YEAR > UB_0]
outliers_0

```

*Summary of column-3: Year*
1.This is a numerical column
2.This represents in which year a particular car is being manufactured
3.There are no outliers



Column -4 : Engine.Fuel.Type

```{r}

#Summary of the column:
summary(dataset_cars$Engine.Fuel.Type)
#finding how many times a word got repeated and finding frequency of the words
table(dataset_cars$Engine.Fuel.Type)
#Converting to lowercase:
dataset_cars$Engine.Fuel.Type <- tolower(dataset_cars$Engine.Fuel.Type)
#Identifying special characters in Engine.Fuel.Type column:
options(max.print = 10000000)
#printing all the special characters
special_characters <- function(column){
   a <- gsub("[a-zA-Z0-9 ]", "", column)
  b <- unlist(strsplit(paste(a, collapse=""), ""))
  b <- unique(b[b != ""])
  cat(b, sep = "")
}

special_characters(dataset_cars$Engine.Fuel.Type)

#removing special characters:
dataset_cars$Engine.Fuel.Type <- gsub(" ","",dataset_cars$Engine.Fuel.Type)
dataset_cars$Engine.Fuel.Type <- gsub("-","",dataset_cars$Engine.Fuel.Type)
dataset_cars$Engine.Fuel.Type <- gsub("/","",dataset_cars$Engine.Fuel.Type)
dataset_cars$Engine.Fuel.Type <- gsub("\\(","",dataset_cars$Engine.Fuel.Type)
dataset_cars$Engine.Fuel.Type <- gsub("\\.","",dataset_cars$Engine.Fuel.Type)
dataset_cars$Engine.Fuel.Type <- gsub("\\)","",dataset_cars$Engine.Fuel.Type)

table(dataset_cars$Engine.Fuel.Type)
```
*Summary od Column-4:*
1.Cleaned and preprocessed the entire column
2.Found special characters in this column and they are [-,/,(,),.]
3.So I removed all the special characters and unwanted spaces
4.The data(column) is preprocessed and cleaned completely.



column-5: Engine.HP

```{r}

summary(dataset_cars$Engine.HP)
#validating data:
is.numeric(dataset_cars$Engine.HP)
#checking null values:
sum(is.na(dataset_cars$Engine.HP))
#Here we found Null values -> Using Mice library to fill null values
#creating a temporary column and filling that column with 1 because mice function needs 2 columns to impute values.
dataset_cars$Engine.HP.2 <- 1 
#imputing values using mice function
imputed.dataset.cars <- mice(dataset_cars[,c("Engine.HP","Engine.HP.2")],
                             m=18,method='pmm',seed=50)
#extracting one fully imputed data set
completed_data_1 <- complete(imputed.dataset.cars, 1)
#Values form completed data set are assigned back to Engine.Engine COLUMN like updating the original column
dataset_cars$Engine.HP <- completed_data_1$Engine.HP
#removing temporary column
dataset_cars$Engine.HP.2 <- NULL 

colnames(dataset_cars) 
#outliers

#treating outliers:
#calculating Interquantile range
IQR_1 <- 300 - 170
LB_1 <- 170 - 1.5 * IQR_1
UB_1 <- 300 + 1.5 * IQR_1
#finding outliers and represemting them using boxplot
outliers_1 <- dataset_cars$Engine.HP[dataset_cars$Engine.HP < LB_1 | 
                                       dataset_cars$Engine.HP > UB_1]
outliers_1
#removing outliers:
#HERE all values lesser than lower bound value will be converted into lower bound vaue
dataset_cars$Engine.HP[dataset_cars$Engine.HP < LB_1] <- LB_1
#HERE all values greater than upper bound value will be converted into upper bound vaue
dataset_cars$Engine.HP[dataset_cars$Engine.HP > UB_1] <- UB_1
#Hence outliers are treated and to represent that thing I used boxplot 
boxplot(dataset_cars$Engine.HP,main="ENGINE.HP_Outlier detection",ylab="values",
        xlab="Engine.HP",col="lightgreen")

```
*summary of column-5:*

1.This is a complete numeric column
2.And there are outliers present in this column
3.I have treated outliers in this column and represented using boxplot
4.I used IQR method to treat outliers
5.There are 62 null values in this column
6.I have treated the null values with mice library and mean imputation technique
7.After this all these steps I ensured that there are no special characters and the column is properly cleaned.


Column-6 : Engine.Cylinders

```{r}
summary(dataset_cars$Engine.Cylinders)
#validating data:
is.numeric(dataset_cars$Engine.Cylinder)
#checking null values:
sum(is.na(dataset_cars$Engine.Cylinder))
#we have null values:
#treating null values:
dataset_cars$Engine.Cylinders2 <- 1 
imputed_dataset2 <- mice(dataset_cars[, c("Engine.Cylinders", 
                      "Engine.Cylinders2")], m = 18, method = 'pmm', seed = 50)

# Extracting one imputed data set
completed_data2 <- complete(imputed_dataset2, 1)

# Updating original column with imputed values
dataset_cars$Engine.Cylinders <- completed_data2$Engine.Cylinders

# Removing the temporary column
dataset_cars$Engine.Cylinders2 <- NULL

colnames(dataset_cars)
#outliers:
#treating outliers:
#calculating Interquantile range
IQR_2 <- 6 - 4
LB_2 <- 4 - 1.5 * IQR_2
UB_2 <- 6 + 1.5 * IQR_2
#finding outliers and represemting them using boxplot
outliers_2 <- dataset_cars$Engine.Cylinders[dataset_cars$Engine.Cylinders < LB_2
                                        | dataset_cars$Engine.Cylinders > UB_2]
outliers_2
#removing outliers:
#HERE all values lesser than lower bound value will be converted into lower bound vaue
dataset_cars$Engine.Cylinders[dataset_cars$Engine.Cylinders < LB_2] <- LB_2
#HERE all values greater than upper bound value will be converted into upper bound vaue
dataset_cars$Engine.Cylinders[dataset_cars$Engine.Cylinders > UB_2] <- UB_2
#representation of boxplot:
boxplot(dataset_cars$Engine.Cylinders,main="ENGINE.Cylinders_Outlier detection",
        ylab="values",xlab="Engine.Cylinders",col="lightgreen")

```
*summary of column-5:*
1.Numeric column
2.Total null values of 25
3.Treated null values with mice library and imputation techniques
4.Outliers are present in this column 
5.Removed outliers and represented with boxplot



column-7: Transmission.Type

```{r}
summary(dataset_cars$Transmission.Type)
#finding how many times a word got repeated and finding frequency of the words
table(dataset_cars$Transmission.Type)
#Converting to lowercase:
dataset_cars$Transmission.Type <- tolower(dataset_cars$Transmission.Type)
#looking for nullvalues:
sum(is.na(dataset_cars$Transmission.Type))
#checking for special characters:
#Using manually created function
special_characters(dataset_cars$Transmission.Type)
#removing special characters from the column
dataset_cars$Transmission.Type <- gsub(" ","",dataset_cars$Transmission.Type)
dataset_cars$Transmission.Type <- gsub("_","",dataset_cars$Transmission.Type)
```
*summary of column-7:*
1.chategorical column
2.Only one special character found and removed that special character
3.No null values

column-8: Driven_Wheels

```{r}
summary(dataset_cars$Driven_Wheels)
#finding how many times a word got repeated and finding frequency of the words
table(dataset_cars$Driven_Wheels)
#Converting to lowercase:
dataset_cars$Driven_Wheels <- tolower(dataset_cars$Driven_Wheels)
#looking for nullvalues:
sum(is.na(dataset_cars$Driven_Wheels))
#checking for special characters:
#Using manually created function
special_characters(dataset_cars$Driven_Wheels)
#removing space between words:
dataset_cars$Driven_Wheels <- gsub(" ","",dataset_cars$Driven_Wheels)

```
*summary of column-8:*
1.categorical column
2.No null values
3.No special characters just unwanted space between words

column-9:Number.of.Doors

```{r}

summary(dataset_cars$Number.of.Doors)
#validating data:
is.numeric(dataset_cars$Number.of.Doors)
#checking null values:
sum(is.na(dataset_cars$Number.of.Doors))
#searching for special characters:
special_characters(dataset_cars$Number.of.Doors)
#filling null values with mice library:
dataset_cars$Number.of.Doors2 <- 1 
imputed_dataset3 <-mice(dataset_cars[, c("Number.of.Doors", "Number.of.Doors2")]
                         , m = 18, method = 'pmm', seed = 50)

# Extracting one imputed data set
completed_data3 <- complete(imputed_dataset3, 1)

# Updating original column with imputed values
dataset_cars$Number.of.Doors <- completed_data3$Number.of.Doors

# Removing the temporary column
dataset_cars$Number.of.Doors2 <- NULL

sum(is.na(dataset_cars$Number.of.Doors))

#outliers:
#In this particular column there's no outliers 
boxplot(dataset_cars$Number.of.Doors,main="NO.OF doors_outlier detection",
        ylab="NO.OF.doors",xlab="Values",col='green')

```
*Summary of column-9:*
1.Numerical column
2.3 null values are there
3.Treated null values with Mice library 
4.No outliers


column-10 : Market.Category

```{r}
summary(dataset_cars$Market.Category)
#finding how many times a word got repeated and finding frequency of the words
#Converting to lowercase:
dataset_cars$Market.Category <- tolower(dataset_cars$Market.Category)
#looking for null values:
sum(is.na(dataset_cars$Market.Category))
#lookng for special characters:
special_characters(dataset_cars$Market.Category)
#removing all the special characters:
dataset_cars$Market.Category <- gsub(" ","",dataset_cars$Market.Category)
dataset_cars$Market.Category <- gsub(",","",dataset_cars$Market.Category)
dataset_cars$Market.Category <- gsub("/","",dataset_cars$Market.Category)
dataset_cars$Market.Category <- gsub("\\-","",dataset_cars$Market.Category)

table(dataset_cars$Market.Category)
```

summary of column-10:
1.chategorical column
2.No null values
3.found some special characters(,/.)
4.So removed all the special characters and unwanted spaces


column-11: Vehicle.Size

```{r}

summary(dataset_cars$Vehicle.Size)
#finding how many times a word got repeated and finding frequency of the words
#Converting to lowercase:
dataset_cars$Vehicle.Size <- tolower(dataset_cars$Vehicle.Size)
#looking for null values:
sum(is.na(dataset_cars$Vehicle.Size))
#lookng for special characters:
special_characters(dataset_cars$Vehicle.Size)
table(dataset_cars$Vehicle.Size)

```
*summary of column-11:*
1.chategorical colum
2.no special characters except spaces
3.removed all the spaces and cleaned
4.No null values


column-12:Vehicle.Style

```{r}

summary(dataset_cars$Vehicle.Style)
#finding how many times a word got repeated and finding frequency of the words
#Converting to lowercase:
dataset_cars$Vehicle.Style <- tolower(dataset_cars$Vehicle.Style)
#looking for null values:
sum(is.na(dataset_cars$Vehicle.Style))
#lookng for special characters:
special_characters(dataset_cars$Vehicle.Style)
#removing space between words:
dataset_cars$Vehicle.Style <- gsub(" ","",dataset_cars$Vehicle.Style)
table(dataset_cars$Vehicle.Style)

```
*summary of column-12:*
1.Categorical column
2.no null values
3.No special characters only spaces are there.


column-13 : highway.MPG

```{r}

summary(dataset_cars$highway.MPG)
#validating data:
is.numeric(dataset_cars$highway.MPG)
#checking null values:
sum(is.na(dataset_cars$highway.MPG))
#searching for special characters:
special_characters(dataset_cars$highway.MPG)
#outliers:
#treating outliers:
#calculating Interquantile range
IQR_3 <- 30 - 22
LB_3 <- 22 - 1.5 * IQR_3
UB_3 <- 30 + 1.5 * IQR_3
#finding outliers and represemting them using boxplot
outliers_3 <- dataset_cars$highway.MPG[dataset_cars$highway.MPG < LB_3 | 
                                         dataset_cars$highway.MPG > UB_3]
outliers_3
#removing outliers:
#HERE all values lesser than lower bound value will be converted into lower bound vaue
dataset_cars$highway.MPG[dataset_cars$highway.MPG < LB_3] <- LB_3
#HERE all values greater than upper bound value will be converted into upper bound vaue
dataset_cars$highway.MPG[dataset_cars$highway.MPG > UB_3] <- UB_3
#Boxplot to prove that there are no outliers:
boxplot(dataset_cars$highway.MPG,main="highway.MPG_Outlier detection",
        ylab="values",xlab="highway.MPG",col="lightgreen")


```

*summary of column-13:*
1.Numerical column
2.There are no null values
3.There are outliers
4.Treated outliers using IQR method.
----------------------------------------------------------------------------------------------------------

column-14: city.mpg

```{r}
summary(dataset_cars$city.mpg)
#validating data:
is.numeric(dataset_cars$city.mpg)
#checking null values:
sum(is.na(dataset_cars$city.mpg))
#searching for special characters:
special_characters(dataset_cars$city.mpg)
#outliers:
#treating outliers:
#calculating Interquantile range
IQR_4 <- 22 - 15
LB_4 <- 15 - 1.5 * IQR_4
UB_4 <- 22 + 1.5 * IQR_4
#finding outliers and represemting them using boxplot
outliers_4 <- dataset_cars$city.mpg[dataset_cars$city.mpg < LB_4 | 
                                      dataset_cars$city.mpg > UB_4]
outliers_4
#removing outliers:
#HERE all values lesser than lower bound value will be converted into lower bound vaue
dataset_cars$city.mpg[dataset_cars$city.mpg < LB_4] <- LB_4
#HERE all values greater than upper bound value will be converted into upper bound vaue
dataset_cars$city.mpg[dataset_cars$city.mpg > UB_4] <- UB_4
#Boxplot to prove that there are no outliers:
boxplot(dataset_cars$city.mpg,main="city.mpg_Outlier detection",ylab="values",
        xlab="city.mpg",col="lightgreen")

```

*summary of column-14:*
1.Numerical column
2.There are no null values
3.There are outliers
4.Treated outliers using IQR method.
----------------------------------------------------------------------------------------------------------------------------

column-15 : Popularity

```{r}

summary(dataset_cars$Popularity)
#validating data:
is.numeric(dataset_cars$Popularity)
#checking null values:
sum(is.na(dataset_cars$Popularity))
#searching for special characters:
special_characters(dataset_cars$Popularity)
#outliers:
#treating outliers:
#calculating Interquantile range
IQR_5 <- 2009 - 549
LB_5 <- 549 - 1.5 * IQR_5
UB_5 <- 2009 + 1.5 * IQR_5
#finding outliers and represemting them using boxplot
outliers_5 <- dataset_cars$Popularity[dataset_cars$Popularity < LB_5 |
                                      dataset_cars$Popularity > UB_5]
outliers_5
#removing outliers:
#HERE all values lesser than lower bound value will be converted into lower bound vaue
dataset_cars$Popularity[dataset_cars$Popularity < LB_5] <- LB_5
#HERE all values greater than upper bound value will be converted into upper bound vaue
dataset_cars$Popularity[dataset_cars$Popularity > UB_5] <- UB_5
#Boxplot to prove that there are no outliers:
boxplot(dataset_cars$Popularity,main="popularity_Outlier detection",ylab="values",
        xlab="popularity",col="lightgreen")

```
*summary of column-15:*
1.Numerical column
2.There are no null values
3.There are outliers
4.Treated outliers using IQR method.
------------------------------------------------------------------------------------------------

column-16: MSRP

```{r}

summary(dataset_cars$MSRP)
#validating data:
is.numeric(dataset_cars$MSRP)
#checking null values:
sum(is.na(dataset_cars$MSRP))
#searching for special characters:
special_characters(dataset_cars$MSRP)
#outliers:
#treating outliers:
#calculating Interquantile range
IQR_6 <- 42146 - 20960
LB_6 <- 20960 - 1.5 * IQR_6
UB_6 <- 42146 + 1.5 * IQR_6
#finding outliers and represemting them using boxplot
outliers_6 <- dataset_cars$MSRP[dataset_cars$MSRP < LB_6 
                                | dataset_cars$MSRP > UB_6]
outliers_6

#removing outliers:
#HERE all values lesser than lower bound value will be converted into lower bound value
dataset_cars$MSRP[dataset_cars$MSRP < LB_6] <- LB_6
#HERE all values greater than upper bound value will be converted into upper bound value
dataset_cars$MSRP[dataset_cars$MSRP > UB_6] <- UB_6
#Boxplot to prove that there are no outliers:
boxplot(dataset_cars$MSRP,main="MSRP_Outlier detection",ylab="values",
        xlab="MSRP",col="lightgreen")

```

*summary of column-16:*
1.Numerical column
2.There are no null values
3.There are outliers
4.Treated outliers using IQR method.

***so the whole process of cleaning is completed.All the columns of the dataset have been cleaned and preprocessed correctly ***

 TASK-2:
 
***Basic Statistical analysis is done in each and everystep of task-1 ***

 Business understanding:
 
 Column-1: MAKE
 ->Categorical column
 -> Highest NO.OF cars manufactured by chevrolet     
 -> Lowest NO.OF cars manufactured by Bugatti
 
 column-2:Model
 ->categorical column
 ->Highest NO.OF car model being sold is Beetle Convertible
 ->There are many models which have sold less than 5.
 
 column-3:Year
 ->Numerical column
 ->Latest car is being manufactured in 2017
 -> Oldest car is manufactured in 1990
 ->No null values
 
 column-4:Engine.Fuel.Type
 ->categorical column
 ->Most NO.OF cars having their engine type as regular unleaded
 ->very few cars are having their engine type as natural gas
 
 column-5:Engine.HP
 ->Numerical column
 ->Mean: 249
 ->Median: 227
 ->Cars having 1001 HP as their engine capacity, which is maximum of all cars
 ->Minimum car HP is 55
 
 column-6:Engine.Cylinders
 ->Numerical column
 ->Mean: 5.86
 ->Median: 6
 ->Maximum no.of Cylinders for a car is 16
 ->And there are cars with no cylinders
 
 column-7:Transmission.Type
 ->Categorical column
 ->Maximum NO.OF cars are Automatic cars which are 6924.
 
 column-8:Driven_wheels:
 ->categorical column
 ->Mainly most NO.OF cars are front wheel drive type cars which are 3996.
 ->There are Less NO.OF four wheel drive  which are 1200.
 
 column-9: NO.Of Doors
 ->Numerical column
 ->Mean:3.43
 ->Median:4
 ->Minimum NO.OF doors per car is 2.
 -> Maximum NO.OF doors per car is 4.
 
 
column-10: Market.Category
->Numerical column
->Mean:3.43
->Median:4
->Mostly Crossover is the market type of 998 cars.

column-11:Vehicle.size
->Categorical column
->3992 cars are being compact.

column-12:Vehicle.Style
->Categorical column
->Most cars are having sedan style structure

column-13:Highway.MPG
->Numerial column
->mean:  26.59
->median: 25.00

column-14:city.mpg
->Numerical column
->mean:19.7
->Median:18

column-15:Popularity
->Numerical column
->Mean:1558
->Median:1385    


column-16:MSRP
->Numerical column
->mean : 40341
->median: 29935  


***completion of business understanding***

**PIECHART**

```{r}

#Piechart of Driven_wheels:
pie_dw <- table(dataset_cars$Driven_Wheels)
pie(pie_dw, main = "Distribution of Driven_Wheels", 
    col = rainbow(length(pie_dw)), radius = 1, cex = 0.9)

```
***bar-chart***

```{r}

#Building bar-graph between Make column
#Using ggplot library for plotting
ggplot(dataset_cars, aes(x = Number.of.Doors , y = Make  , fill = Make)) +
  geom_bar(stat = "identity") + 
  labs(title = "Car Makes Distribution", x = "Car Make", y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1.2))

```


***histogram***


```{r}
#histogram of Year column:
ggplot(dataset_cars, aes(x = Year)) +
  geom_histogram(binwidth = 1, fill = "yellow", color = "black") +
  labs(title = "Distribution of Year", x = "Year", y = "Frequency") +
  theme_minimal()

```


***Scatter plot***


```{r}

# Scatterplot of highway.MPG vs city.mpg
plot(dataset_cars$highway.MPG,dataset_cars$city.mpg, 
     main="highway.MPG vs city.mpg",xlab="highway.MPG",ylab="city.mpg",
     pch=19,col="blue")

```

*TASK - 3*

Analysis on price variable:


```{r}
# 4A :
#histogram of price variable:
hist(dataset_cars$MSRP,main="histogram of Price column",xlab="PRICE",
              ylab="FREQUENCY",col="yellow",border="black")
``` 
***Explanation and summary of Histogram***

- From the above histogram of price(MSRP) column:
- More than 1250 cars price ranges from 0 - 5000
- Very few cars price ranges from 5000 - 10000
- More than 100 cars price ranges from 10000 - 15000
- More than 750 cars price ranges from 15000 - 20000
- More than 1200 cars price ranges from 20000 - 25000
- More than 1250 cars price ranges from 25000 - 30000
- More than 1100 cars price ranges from 30000 - 35000
- More than 750 cars price ranges from 35000 - 40000
- More than 600 cars price ranges from 40000 - 45000
- More than 400 cars price ranges from 45000 - 50000
- More than 200 cars price ranges from 50000 - 55000
- More than 100 cars price ranges from 55000 - 60000
- More than 120 cars price ranges from 60000 - 65000
- More than 100 cars price ranges from 65000 - 70000
- More than 800 cars price are from 70000 and above.



***Calculation of mean median variance***

```{r}

#calculating mean:
mean(dataset_cars$MSRP)

#median of MSRP
median(dataset_cars$MSRP)

#varience of price:
var(dataset_cars$MSRP)
```


```{r}
# 4b:

#Group cars by price ranges
price_groups <- cut(dataset_cars$MSRP, breaks = c(0, 20000, 50000, 100000, Inf),
                    labels = c("Low(<20)", "Medium (20K-50K)", "High(50K-100K)",
                               "Luxury(>100K)"), include.lowest = TRUE)
group_summary <- table(price_groups)
print(group_summary)
```


```{r}
# 4c:

# Boxplot for different car types
vehicle_style_01<- names(sort(table(dataset_cars$Vehicle.Style),
                              decreasing = TRUE))[1:5]
vehicle_style_01<-dataset_cars[dataset_cars$Vehicle.Style %in% vehicle_style_01,]


par(mar = c(11,5.5,2,1) +0.1)
par(mgp = c(3, 1, 0))  

boxplot(MSRP ~ Vehicle.Style, data = vehicle_style_01, 
        main = "Price for Top 5 Vehicle Styles", xlab = " ", 
        ylab = " ", col = "yellow", las = 2, ylim = c(0, 200000))
mtext("Style of the vehicle", side = 1, line = 6, cex = 1)
mtext("MSRP(Price)", side = 2, line = 4, cex = 1, las = 3)

```



```{r}
#4[d] -> Correlation with MSRP (numeric variables only)
numeric_variables <- sapply(dataset_cars, is.numeric)
correlation_matrix <- cor(dataset_cars[, numeric_variables], use = "complete.obs")
msrp_correlation <- correlation_matrix["MSRP", ]
msrp_correlation <- msrp_correlation[!names(msrp_correlation) %in% "MSRP"]
sorted_correlation <- sort(abs(msrp_correlation), decreasing = TRUE)
top_3_correlation <- sorted_correlation[1:3]

print(msrp_correlation)
print("Top 3 correlated variables with MSRP -> Price:")
print(top_3_correlation)

```

***Brand affecting the popularity and the price***

```{r}

#5:
#Summarize MSRP(Price) by Make:
brand_summary_01 <- aggregate(MSRP ~ Make, data = dataset_cars, FUN = function(x)
  c(mean = mean(x), median = median(x)))
brand_summary_01 <- brand_summary_01[order(-brand_summary_01$MSRP[, "mean"]), ]
print(head(brand_summary_01, 10))

# Representing MSRP column using boxplot:
# Boxplot of MSRP(Price) for top 10 Makes:

top10.makes <- brand_summary_01$Make[1:10]
top10.makes <- dataset_cars[dataset_cars$Make %in% top10.makes, ]  
par(mar = c(11,5.5,2,1) + 0.1)
par(mgp = c(3, 1, 0))

boxplot(MSRP ~ Make, data = top10.makes, main ="Price(MSRP) of Top 10 Makes", 
        xlab = "", ylab = "", col= "lightgreen", las = 2, 
        cex.names = 0.9, ylim = c(0,200000))

mtext("Make", side = 1, line = 6, cex = 1)
mtext("MSRP", side = 2, line = 4, cex = 1, las = 3)

```

  ***THANK YOU***


